Synchrotron X-Ray Rims in Tycho's Supernova Remnant are Energy Dependent

### 2014-09-27 manuscript/paper-tycho.tex
Kepler/CasA pipeline, fits; paper cleanup
Merge projection splitting + box convert in specextract pipeline
Additional Cas A regions.  FWHMs for new Kepler regions.
Remove old tables/plots from manuscript

Evidence for acceleration: spectral softening \citep{cassam-chenai2007}, CD-FS
gap \citep{warren2005}, multi-wavelength observations \citep{acero2010,
acciari2011, morlino2012} for SN 1006 and Tycho; Tycho's tiger stripes
\citep{eriksen2011}. These further support the notion that Tycho is actively
accelerating hadrons and leptons.
Various questions to consider (or answer).
Why the diversity of filament structure? How does diffusion really work?  How
much energy is going into cosmic rays / how efficient is the acceleration
process?  All sorts of things... (one interesting paper: \citet{caprioli2013}).
\note[Aaron]{this para may not be necessary}

@@

We selected 13 regions in 5 distinct filaments around Tycho's shock
(Figure~\ref{fig:snr}) based on the following criteria: (1) filaments should be
clear of spatial plumes of thermal ejecta in \Chandra images, which rules out,
e.g., areas of strong thermal emission on Tycho's eastern limb; (2) filaments
should be singular and localized; multiple filaments should either not
overlap or completely overlap (rules out parts of NE limb); (3) filaments
should have clear FWHMs in most energy band; i.e., a peak should be evident
above the background signal or downstream thermal emission (rules out some
faint southern wisps) \note[Aaron]{wishy-washy qualitative}.  We grouped
regions into filaments by visual inspection of the remnant.

% NOTE generate final, higher quality figure + newer figure for regions-5
\begin{figure}
    \centering
    \plotone{figures/f0-snr-reg5-inv.png}
    \caption{RGB image of Tycho with region selections overlaid.  Bands are
    0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).
    \note[Aaron]{Could show the spectrum extraction regions on here too?
    Maybe too cluttered.  Figure colors inverted to save ink.}}
    \label{fig:snr}
\end{figure}

@@

% NOTE be consistent in usage of word 'sections'
We extracted spectra for all regions to check that rim width measurements are
not affected by contaminating thermal line emission.  In each region, we
selected upstream and downstream sections from which to extract spectra.  The
upstream section is the smallest sub-region that contains the measured FWHM
bounds from all energy bands (see Section~\ref{sec:fwhms}); i.e., this region
captures the thin rim.  The downstream section extends from the back of the rim
(end of the upstream section) to the rim model fit's (equation~\eqref{eq:prof})
downstream limit.  We merged spectra from all \Chandra ObsIDs and grouped the
data into bins of $\geq 15$ counts.  Figure~\ref{fig:spec} plots an example
profile (4.5--7 keV) with the downstream and upstream sections highlighted; the
dividing line is set at the \textbf{???--???} energy band's downstream FWHM
limit here (see Figure~\ref{fig:profiles}).
We extracted background spectra from circular regions (radius $\abt 30\arcsec$)
around the remnant's exterior; each thin rim region was matched to the closest
background region's spectrum.
% Background region size is valid for data-tycho/bkg-2/ selections

% NOTE must match fig:profiles!  Keep region numbers updated and
% consistent.  Currently, region 1 of data-tycho/regions-4/ selections
\begin{figure*}
    \plotone{figures/fig-specs.pdf}
    \caption{Spectra and fits from Region 1. Left: $4.5$--$7 \unit{keV}$
    profile with highlighted downstream (blue) and upstream (grey) sections.
    Intensity units are arbitrary (a.u.).  Middle: downstream spectrum with
    absorbed power law fit; Si and S lines at $1.85$, $2.45 \unit{keV}$ are
    clearly visible.  Right: upstream spectrum with absorbed power law fit
    shows that the filament is likely free of thermal line emission.}
    \label{fig:spec}
\end{figure*}

We fit $0.5$--$7 \unit{keV}$ spectra from each region's upstream and downstream
sections to an absorbed power law model (XSPEC 12.8.1, \texttt{phabs*po}) with
spectral index $n$, hydrogen column density $n_H$, and a normalization as free
parameters.  Table~\ref{tab:spec} lists best fit parameters and reduced
$\chi^2$ values for all regions.  Spectra from thin rims (upstream sections)
are well-fit by the power law model alone; spectra from sections downstream of
the rims are generally poorly-fit, reflecting thermal contamination primarily
from \ion{Si}{13} and \ion{S}{15} He$\alpha$ line emission at $1.85$ and $2.45$
keV.  Although most downstream spectra show thermal contamination, some regions
(e.g., 4, 5 on Tycho's northwestern limb) are reasonably well-fit downstream by
an absorbed power law as well due to a larger distance between the forward
shock and contact discontinuity.
Upstream spectra fits spectral indices between $2.5$--$3$ are consistent with
synchrotron radiation spectra for a steep power-law electron distribution
\citep{vink2012}. Best fit column densities from
upstream spectra between $0.61$--$0.78 \times 10^{22} \unit{cm^{-2}}$ are
consistent with previous fits to Tycho's nonthermal rims \citep{hwang2002} but
somewhat larger than $\abt0.35$--$0.5 \times 10^{22} \unit{cm^{-2}}$ from radio
and optical??  HI absorption line measurements \citep{black1984, albinson1986,
kothes2004}.
% NOTE keep stated ranges of spectral indices, column densities up to date

% NOTE keep region numbers updated (in discussion of good/bad downstream spectra)
\begin{table}
    \scriptsize
    \centering
    \caption{Region spectra fit parameters\label{tab:spec}}
    \input{tab-spec.tex}
    \tablecomments{Absorbed power law fit parameters are photon index $\Gamma$
    and hydrogen column density $\mathrm{N_H}$.}
\end{table}

@@

\begin{figure*}[ht]
    \plotone{figures/fig-profiles.pdf}
    \caption{Best fit profiles with measured FWHMs demarcated for each energy
             band in Region 1.  Red data points are excluded from profile fit
             domain.}
    \label{fig:profiles}
\end{figure*}

% NOTE keep blacklisted/excluded region numbers updated!
From the fitted profiles we extracted a full width at half maximum (FWHM) for
each region and each energy band.  We could not resolve a FWHM in regions 2, 6,
and 8 at 0.7--1 keV (Table~\ref{tab:fwhms}); in these regions, either
the downstream FWHM bound would extend outside the fit domain or we could not
find an acceptable fit to equation~\eqref{eq:prof}.  We were able to resolve
FWHMs for all regions at higher energy bands (1--7 keV).
% NOTE review the FWHM rejection criterion, on the final set of regions

@@

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Measured FWHMs for all regions.\label{tab:fwhms}}
    \input{tab-fwhms-reg5.tex}
\end{table*}

@@

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filaments 1--3.
    \label{tab:fits}}
    \input{tab-fits-all-reg5.tex}
\end{table*}

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filaments 4--5.
    \label{tab:fits-pt2}}
    \input{tab-fits-all-reg5-pt2.tex}
\end{table*}

@@

\begin{table*}[ht]
    \scriptsize
    \centering
    \caption{Filament averages of best-fit parameters for constituent regions
    \label{tab:fits-avg}}
    \input{tab-fits-par-avg-reg5.tex}
\end{table*}

>>>>

Evidence for acceleration: basically, see \citet{morlino2012} for a nice review
of evidence in Tycho.  No need to cite everything here.

spectral softening \citep{cassam-chenai2007}, CD-FS gap \citep{warren2005},
multi-wavelength observations \citep{acero2010, acciari2011, morlino2012} for
SN 1006 and Tycho; Tycho's tiger stripes \citep{eriksen2011}. These further
support the notion that Tycho is actively accelerating hadrons and leptons.
Various questions to consider (or answer).  Why the diversity of filament
structure? How does diffusion really work?  How much energy is going into
cosmic rays / how efficient is the acceleration process?  All sorts of
things... (one interesting paper: \citet{caprioli2013}).
\note[Aaron]{this para may not be necessary}

@@

% KEEP REGION NUMBER UPDATED
We selected 22 regions in 5 distinct filaments around Tycho's shock
(Figure~\ref{fig:snr}) based on the following criteria: (1) filaments should be
clear of spatial plumes of thermal ejecta in \Chandra images, which rules out,
e.g., areas of strong thermal emission on Tycho's eastern limb; (2) filaments
should be singular and localized; multiple filaments should either not
overlap or completely overlap (rules out parts of NE limb); (3) filaments
peaks should be evident above the background signal or downstream thermal
emission (rules out some faint southern wisps), although we selected some
regions with poor quality peaks in the lowest energy band (0.7--1 keV) so long
as peaks in all higher energy bands were acceptable and well-fit.
\note[Aaron]{vague, weakly qualitative}
We grouped regions into filaments by visual inspection of the remnant.
Within each filament, we chose region widths to get a similar number of counts
at the thin rim, in each region.
We attempted to balance signal-to-noise of each profile against having multiple
selections for each filament (a signal-to-noise further downstream, in a
sense); within each filament we tried to keep the number of counts in each
region roughly consistent... \note[Aaron]{wording for latter half of para}

% NOTE generate final, higher quality figure + newer figure for regions-5
\begin{figure}
    \centering
    \plotone{figures/f0-snr.png}
    \caption{RGB image of Tycho with region selections overlaid.  Bands are
    0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).
    \note[Aaron]{Could show the spectrum extraction regions on here too?
    Maybe too cluttered.  Figure colors inverted to save ink.}}
    \label{fig:snr}
\end{figure}

@@

% NOTE be consistent in usage of word 'sections'
We extracted spectra for all regions to check that rim width measurements are
not affected by contaminating thermal line emission.  In each region, we
selected upstream and downstream sections from which to extract spectra.  The
upstream section is the smallest sub-region that contains the measured FWHM
bounds from all energy bands (see Section~\ref{sec:fwhms}); i.e., this region
captures the thin rim.  The downstream section extends from the back of the rim
(end of the upstream section) to the rim model fit's (equation~\eqref{eq:prof})
downstream limit.  We merged spectra from all \Chandra ObsIDs and grouped the
data into bins of $\geq 15$ counts.  Figure~\ref{fig:spec} plots an example
profile (4.5--7 keV) with the downstream and upstream sections highlighted.
We extracted background spectra from circular regions (radius $\abt 30\arcsec$)
around the remnant's exterior; each thin rim region was matched to the closest
background region's spectrum.
% Background region size is valid for data-tycho/bkg-2/ selections

% NOTE must match fig:profiles!  Keep region numbers updated and
% consistent.  Currently, region 1 of data-tycho/regions-4/ selections
\begin{figure*}
    \plotone{figures/fig-specs_01.pdf}
    \plotone{figures/fig-specs_16.pdf}
    \caption{Spectra and fits from Regions 1 (top) and 16 (bottom).
    Left: $4.5$--$7 \unit{keV}$
    profile with highlighted downstream (blue) and upstream (grey) sections.
    Intensity in arbitrary units (a.u.).  Middle: downstream spectrum with
    absorbed power law fit; Si and S lines at $1.85$, $2.45 \unit{keV}$ are
    clearly visible.  Right: upstream spectrum with absorbed power law fit
    shows that each region is likely free of thermal line emission.}
    \label{fig:spec}
\end{figure*}

We fit $0.5$--$7 \unit{keV}$ spectra from each region's upstream and downstream
sections to an absorbed power law model (XSPEC 12.8.1, \texttt{phabs*po}) with
spectral index $n$, hydrogen column density $n_H$, and a normalization as free
parameters.  Table~\ref{tab:spec} lists best fit parameters and reduced
$\chi^2$ values for all regions.  Spectra from thin rims (upstream sections)
are well-fit by the power law model alone; spectra from sections downstream of
the rims are generally poorly-fit, reflecting thermal contamination primarily
from \ion{Si}{13} and \ion{S}{15} He$\alpha$ line emission at $1.85$ and $2.45$
keV.  Although most downstream spectra show thermal contamination, some regions
(e.g., 4, 5 on Tycho's northwestern limb) are reasonably well-fit downstream by
an absorbed power law as well due to a larger distance between the forward
shock and contact discontinuity.
Upstream spectra fits spectral indices between $2.5$--$3$ are consistent with
synchrotron radiation spectra for a steep power-law electron distribution
\citep{vink2012}. Best fit column densities from
upstream spectra between $0.61$--$0.78 \times 10^{22} \unit{cm^{-2}}$ are
consistent with previous fits to Tycho's nonthermal rims \citep{hwang2002} but
somewhat larger than $\abt0.35$--$0.5 \times 10^{22} \unit{cm^{-2}}$ from radio
and optical??  HI absorption line measurements \citep{black1984, albinson1986,
kothes2004}.
% NOTE keep stated ranges of spectral indices, column densities up to date
% NOTE keep example region numbers up to date

% NOTE keep region numbers updated (in discussion of good/bad downstream spectra)
\begin{table}
    \scriptsize
    \centering
    \caption{Region spectra fit parameters\label{tab:spec}}
    \input{tab-spec.tex}
\end{table}

@@

\begin{figure*}[ht]
    \plotone{figures/fig-profiles_01.pdf}
    \plotone{figures/fig-profiles_16.pdf}
    \caption{Best fit profiles with measured FWHMs demarcated for each energy
        band in Region 1 (top) and Region 16 (bottom).  Energy bands increase
        from left to right.  Data points in red were excluded from profile
        fitting domains as described in text.}
    \label{fig:profiles}
\end{figure*}

% NOTE keep blacklisted/excluded region numbers updated!
From the fitted profiles we extracted a full width at half maximum (FWHM) for
each region and each energy band.  We could not resolve a FWHM in 7 of 22
regions at 0.7--1 keV (Table~\ref{tab:fwhms}); in these regions, either
the downstream FWHM bound would extend outside the fit domain or we could not
find an acceptable fit to equation~\eqref{eq:prof}.  We were able to resolve
FWHMs for all regions at higher energy bands (1--7 keV).
% NOTE review the FWHM rejection criterion, on the final set of regions

@@

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Measured FWHMs for all regions.\label{tab:fwhms}}
    \input{tab-fwhms.tex}
\end{table*}

@@

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filaments 1--3.
    \label{tab:fits}}
    \input{tab-fits-all.tex}
\end{table*}

\begin{table*}[ht]
    \tiny
    \centering
    \caption{Full model best fits for individual regions, Filaments 4--5.
    \label{tab:fits-pt2}}
    \input{tab-fits-all-pt2.tex}
\end{table*}

@@

\begin{table*}[ht]
    \scriptsize
    \centering
    \caption{Filament averages of best-fit parameters for constituent regions
    \label{tab:fits-avg}}
    \input{tab-fits-par-avg.tex}
\end{table*}
